Technical Interview

### 2017-09-06 tex/paper/paper.tex
final paper v1.1

\subsection{General Notes}
We use efficient data structures as appropriate and we modelise the problem using
the whole power of Haskell. 
We use \emph{newtype} as much as possible as it allows design at a 
seemless cost.

@@

\subsection{Uniform Engine Module}
\label{ssec:uniform.impl}
The \emph{UniformRNG} ADT abstracts a choice of uniform pseudo random 
number generator, Mersenne Twister and Ecuyer. We could seemlessly
add new generator, if needs be. We would need to extend the standard
Haskell \emph{RandomGen} class. 
They can generate un-bounded pseudo random numbers.

@@

\subsection{Weighted Engine Module}
\label{ssec:weighted.impl}
The \emph{ProbaWPEngine} ADT allows to computed weighted pseudo
random numbers. It order to achieve this, it need to build and keep
thoughout the computations the \emph{EngineParams} ADT. 
This latter encapsulates the \emph{PDF}, \emph{CDF} and \emph{InvCDF}
ADTs. They do not change thoughout the computations and they are kept 
in memory for efficiency. We pass them thought the computations
using the \emph{Reader} monad. We also use the \emph{State} monad
to pass on the state of the bounded uniform engine.
It also instanciates the \emph{ProbaEngine} class.

\subsection{Statistics Engine Module}
\label{ssec:stats.impl}
This \emph{StatEngine} type role is
to compute the statistics of a series of random numbers generated
by either the \emph{ProbaUniEngine} or the \emph{ProbaWPEngine}.

@@

\subsection{CLI Module}
Using the \emph{Query} ADT we modelise a query from the command line
dynamically. This latter is processed by the \emph{run} function to 
produce a  \emph{Result} ADT.

@@

\subsection{Testing Module}
We modelise a \emph{TestSuite} as a list of \emph{Test}.
Each \emph{Test} can be of different type: QuickCheck type - 
using a pseudo random data, 
an IO type - an action that need to be run, e.g. performance tests, 
and a Pure kind of test, for pure computation action.

%-------------------------------------------------------------------------------
\section{Qualitative Analysis}
\label{sec:quality}
We realise that we maybe be beyond the scope of the initial
question, which was to implement a \emph{nextNum} function.
However how to test it efficiently ? It is a very hard question.
This led us to look at the distribution of the results to ensure
the quality of the numbers generated.

@@

\subsection{Analytics Testing} 
These are the unit tests for the Analytics only.
\begin{description}
	\item [CDF creation]
		Given a 1000 pseudo random PDF, we ensure 
		that the last pillar of the CDF is 1.
		As the CDF is build incrementally, this gives a good sense that
		the CDF construction has been correct.
	\item [InverseCDF, CDF and PDF pillars check]
		Given a 1000 pseudo random PDF, we ensure that
		the pillars on the x axis of the PDF and CDF are the pillars on the y
		axis of the Inverse CDF.
	\item [PDF Stable]
		Given a 1000 pseudo random PDF pillars, 
		we ensure that we get the same PDF given the reverse input.
	\item [PDF consistency]
		Given a 1000 pseudo random PDF, 
		we ensure that the sum of the PDF pillars probabilities are 
		equal to 1. We indeed modelise the non selected element
		only in the CDF.
	\item [Inverse CDF valid]
		Given a 1000 pseudo random PDF,
		we build the CDF and we ensure that for a random number of
		x values within 0 and 1 we always recover a x pillars of the PDF.
	\item [PDF failure]
		We test the expecture failure of the PDF construction. 
		We can't have negative values. We need at least one valid pillar.
		The pillars with 0 probabilities are discarded.
		And the sum of the PDF pillars probabilities must be 
		equal to one.
	\item [Inverse CDF Failure]
		We check that the inverse CDF function raise an
		exception if the x input is outside the correct bounds, 
		$\left[0, 1\right]$.
	\item [Mean and Standard Deviation Failure]
		These functions raise an exception for empty lists.
		We test it automatically.
	\item [PDF and Inverse CDF Non Regression]
		We ensure without randomness that given a set of pillars,
		we recover the expected Inverse CDF and PDF.
	\item [Statistics Non Regression]
		We ensure without randomness that given a set of pillars,
		we recover the expected statistics used in the post processing
		of the Stats engines.
\end{description}
\subsection{Uniform Engine Testing}

@@

\begin{description}
	\item [Mean and Standard Deviation]
		Given a large number of random number generated for
		both Ecuyer and Mersenne, i.e. $100000$.
		We ensure that the mean and the standard deviation 
		of the distributions are still the same. Run 10 times.
		We ensure that the uniform random number engines keep
		the same properties.
	\item [Bounds]
		Test ensuring that the uniform pseudo random generators
		generate probabilities with the $[0, 1]$ bound.
		Run $10$ times for $1000000$ random numbers.
	\item [Uniform]
		Ensure that the ditributions generated are uniform,
		by making sure that we match a flat distribution at 0.001
		precision on the standard deviation of the difference
		between the PDFs.
		Run 10 times for $100000$ random numbers.
	\item [Performance]
		Ensure that the pseudo random numbers keep at
		least the same performance.
		Run 10 times for $100000$ random numbers.
	\item [Engine Fail]
		Ensure that the construction of the engine 
		handles all the expected failure.
\end{description}

\subsection{Weighted Engine Testing} 
These are the unit tests for the weighted pseudo random engine.
All tests are run using the two available uniform pseudo random
generators. 
These are like the integration tests as they use most of the library.
\begin{description}
	\item [Distribution check]
		We check that we recover the PDF distribution from
		a sample of 1000000 random numbers at $0.1\%$ precision
		on the standard deviation of the differences with the 
		generated PDF.
		We run this for 10 pseudo random PDF.
	\item [Engine Fail]
		Ensure that the construction of the engine 
		handles all the expected failure.
	\item [Performance]
		Ensure that the engine keep at least the same performance.
		Run 10 times for $100000$ random  numbers	
\end{description}

\subsection{CLI Testing}
We test that the API produce what is expected.
\begin{description}
	\item [Query]
	We test that given an argument list, we produce the expected
	Query.
	\item[Run]
	We test that for given a Query, we produce the expected Result.
\end{description}

%-------------------------------------------------------------------------------
\section{Performance Analysis}
\label{sec:perf}

@@

%-------------------------------------------------------------------------------
\section{Possible Extensions}
\label{sec:ext}
Here is a list of possible extension we could add.
\begin{description}
	\item [Testing]
	We could run the tests on different environments to ensure the portability of 
	the code.
	\item [UI] 
	A nice extension would be to add a Graphic User Interface,
	using the library 
	\href{https://hackage.haskell.org/package/threepenny-gui}{threepenny-gui}
	to create a portable Web application.

@@

\chapter{Code} \label{apx:code}
We provide below the code. We use Vim with the extension 
\href{https://github.com/begriffs/haskell-vim-now}{haskell-vim-now}
\end{document}

>>>>

\subsection{General Notes}
We use efficient data structures as appropriate and we modelise the problem using
the whole power of Haskell. 
The types are key, c.f. the code at \autoref{fig:core.types1}, \autoref{fig:core.types2}
and \autoref{fig:core.types3}.

We use \emph{newtype} as much as possible as it allows design at a 
seemless cost.

@@

The relevant code is in \autoref{fig:core.analytics}.
\subsection{Uniform Engine Module}
\label{ssec:uniform.impl}
The \emph{UniformRNG} ADT abstracts a choice of uniform pseudo random 
number generator, Mersenne Twister and Ecuyer. We could seemlessly
add new generator, if needs be. We would need to extend the standard
Haskell \emph{RandomGen} class. 
They can generate un-bounded pseudo random numbers.

@@

The relevant code is in \autoref{fig:core.u}.
\subsection{Weighted Engine Module}
\label{ssec:weighted.impl}
The \emph{ProbaWPEngine} ADT allows to computed weighted pseudo
random numbers. It order to achieve this, it need to build and keep
thoughout the computations the \emph{EngineParams} ADT. 
This latter encapsulates the \emph{PDF}, \emph{CDF} and \emph{InvCDF}
ADTs. They do not change thoughout the computations and they are kept 
in memory for efficiency. We pass them thought the computations
using the \emph{Reader} monad. We also use the \emph{State} monad
to pass on the state of the bounded uniform engine.
It also instanciates the \emph{ProbaEngine} class.

The relevant code is in \autoref{fig:core.we}.
\subsection{Statistics Engine Module}
\label{ssec:stats.impl}
This \emph{StatEngine} type role is
to compute the statistics of a series of random numbers generated
by either the \emph{ProbaUniEngine} or the \emph{ProbaWPEngine}.

@@

The relevant code is in \autoref{fig:core.se}.
\subsection{CLI Module}
Using the \emph{Query} ADT we modelise a query from the command line
dynamically. This latter is processed by the \emph{run} function to 
produce a  \emph{Result} ADT.

@@

The relevant code is in \autoref{fig:core.cli1} and \autoref{fig:core.cli2}.
\subsection{Testing Module}
We modelise a \emph{TestSuite} as a list of \emph{Test}.
Each \emph{Test} can be of different type: QuickCheck type - 
using a pseudo random data, 
an IO type - an action that need to be run, e.g. performance tests, 
and a Pure kind of test, for pure computation action.

The relevant code is in \autoref{fig:test.qc1} and \autoref{fig:test.qc2}.
%-------------------------------------------------------------------------------
\section{Qualitative Analysis}
\label{sec:quality}
We realise that we maybe be beyond the scope of the initial
question, which was to implement a \emph{nextNum} function.
However how to test it efficiently ? It is a very hard question.
This led us to look at the distribution of the results to ensure
the quality of the numbers generated.

@@

\subsection{Analytics Testing} 
These are the unit tests for the Analytics only.
\begin{description}
	\item [CDF creation]
		Given a 1000 pseudo random PDF, we ensure 
		that the last pillar of the CDF is 1.
		As the CDF is build incrementally, this gives a good sense that
		the CDF construction has been correct.
	\item [InverseCDF, CDF and PDF pillars check]
		Given a 1000 pseudo random PDF, we ensure that
		the pillars on the x axis of the PDF and CDF are the pillars on the y
		axis of the Inverse CDF.
	\item [PDF Stable]
		Given a 1000 pseudo random PDF pillars, 
		we ensure that we get the same PDF given the reverse input.
	\item [PDF consistency]
		Given a 1000 pseudo random PDF, 
		we ensure that the sum of the PDF pillars probabilities are 
		equal to 1. We indeed modelise the non selected element
		only in the CDF.
	\item [Inverse CDF valid]
		Given a 1000 pseudo random PDF,
		we build the CDF and we ensure that for a random number of
		x values within 0 and 1 we always recover a x pillars of the PDF.
	\item [PDF failure]
		We test the expecture failure of the PDF construction. 
		We can't have negative values. We need at least one valid pillar.
		The pillars with 0 probabilities are discarded.
		And the sum of the PDF pillars probabilities must be 
		equal to one.
	\item [Inverse CDF Failure]
		We check that the inverse CDF function raise an
		exception if the x input is outside the correct bounds, 
		$\left[0, 1\right]$.
	\item [Mean and Standard Deviation Failure]
		These functions raise an exception for empty lists.
		We test it automatically.
	\item [PDF and Inverse CDF Non Regression]
		We ensure without randomness that given a set of pillars,
		we recover the expected Inverse CDF and PDF.
	\item [Statistics Non Regression]
		We ensure without randomness that given a set of pillars,
		we recover the expected statistics used in the post processing
		of the Stats engines.
\end{description}
The relevant code is available in \autoref{fig:test.a1}, \autoref{fig:test.a2}
and \autoref{fig:test.a3}.

\subsection{Uniform Engine Testing}

@@

\begin{description}
	\item [Mean and Standard Deviation]
		Given a large number of random number generated for
		both Ecuyer and Mersenne, i.e. $100000$.
		We ensure that the mean and the standard deviation 
		of the distributions are still the same. Run 10 times.
		We ensure that the uniform random number engines keep
		the same properties.
	\item [Bounds]
		Test ensuring that the uniform pseudo random generators
		generate probabilities with the $[0, 1]$ bound.
		Run $10$ times for $1000000$ random numbers.
	\item [Uniform]
		Ensure that the ditributions generated are uniform,
		by making sure that we match a flat distribution at 0.001
		precision on the standard deviation of the difference
		between the PDFs.
		Run 10 times for $100000$ random numbers.
	\item [Performance]
		Ensure that the pseudo random numbers keep at
		least the same performance.
		Run 10 times for $100000$ random numbers.
	\item [Engine Fail]
		Ensure that the construction of the engine 
		handles all the expected failure.
\end{description}
The relevant code is in \autoref{fig:test.u1} and \autoref{fig:test.u2}.
\subsection{Weighted Engine Testing} 
These are the unit tests for the weighted pseudo random engine.
All tests are run using the two available uniform pseudo random
generators. 
These are like the integration tests as they use most of the library.
\begin{description}
	\item [Distribution check]
		We check that we recover the PDF distribution from
		a sample of 1000000 random numbers at $0.1\%$ precision
		on the standard deviation of the differences with the 
		generated PDF.
		We run this for 10 pseudo random PDF.
	\item [Engine Fail]
		Ensure that the construction of the engine 
		handles all the expected failure.
	\item [Performance]
		Ensure that the engine keep at least the same performance.
		Run 10 times for $100000$ random  numbers	
\end{description}	
The relevant code is in \autoref{fig:test.w}.
\subsection{CLI Testing}
We test that the API produce what is expected.
\begin{description}
	\item [Query]
	We test that given an argument list, we produce the expected
	Query.
	\item[Run]
	We test that for given a Query, we produce the expected Result.
\end{description}
The relevant code is in \autoref{fig:test.cli}.
%-------------------------------------------------------------------------------
\section{Performance Analysis}
\label{sec:perf}

@@

%-------------------------------------------------------------------------------
\section{Possible Extensions}
\label{sec:ext}
Here is a list of possible extension we could add.
\begin{description}
	\item [Testing]
	We could run the tests on different environments to ensure the portability of 
	the code.

	We could also automate the performance testing.
	\item [UI] 
	A nice extension would be to add a Graphic User Interface,
	using the library 
	\href{https://hackage.haskell.org/package/threepenny-gui}{threepenny-gui}
	to create a portable Web application.

@@

\chapter{Code} \label{apx:code}
We provide below the code. We use Vim with the extension 
\href{https://github.com/begriffs/haskell-vim-now}{haskell-vim-now}.
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{img/code-run.png}
\caption{Code: ManAhl.Run}
\label{fig:core.run}
\end{figure}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{img/code-types-1.png}
\caption{Code: ManAhl.Core.Types - page 1}
\label{fig:core.types1}
\end{figure}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{img/code-types-2.png}
\caption{Code: ManAhl.Core.Types - page 2}
\label{fig:core.types2}
\end{figure}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{img/code-types-3.png}
\caption{Code: ManAhl.Core.Types - page 3}
\label{fig:core.types3}
\end{figure}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{img/code-ue.png}
\caption{Code: ManAhl.Core.UniformEngine}
\label{fig:core.u}
\end{figure}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{img/code-we.png}
\caption{Code: ManAhl.Core.WeightedEngine}
\label{fig:core.we}
\end{figure}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{img/code-se.png}
\caption{Code: ManAhl.Core.StatEngine}
\label{fig:core.se}
\end{figure}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{img/code-analytics.png}
\caption{Code: ManAhl.Core.Analytics}
\label{fig:core.analytics}
\end{figure}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{img/code-cli-1.png}
\caption{Code: ManAhl.CLI - page 1}
\label{fig:core.cli1}
\end{figure}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{img/code-cli-2.png}
\caption{Code: ManAhl.CLI - page 2}
\label{fig:core.cli2}
\end{figure}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{img/code-test-run.png}
\caption{Code: Test.ManAhl.Run}
\label{fig:test.run}
\end{figure}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{img/code-test-quickcheck-1.png}
\caption{Code: Test.ManAhl.QuickCheck - page 1}
\label{fig:test.qc1}
\end{figure}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{img/code-test-quickcheck-2.png}
\caption{Code: Test.ManAhl.QuickCheck - page 2}
\label{fig:test.qc2}
\end{figure}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{img/code-test-uni-1.png}
\caption{Code: Test.ManAhl.Core.UniformEngine - page 1}
\label{fig:test.u1}
\end{figure}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{img/code-test-uni-2.png}
\caption{Code: Test.ManAhl.Core.UniformEngine - page 2}
\label{fig:test.u2}
\end{figure}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{img/code-test-w.png}
\caption{Code: Test.ManAhl.Core.WeightedEngine }
\label{fig:test.w}
\end{figure}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{img/code-test-analytics-1.png}
\caption{Code: Test.ManAhl.Core.Analytics - page 1}
\label{fig:test.a1}
\end{figure}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{img/code-test-analytics-2.png}
\caption{Code: Test.ManAhl.Core.Analytics - page 2}
\label{fig:test.a2}
\end{figure}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{img/code-test-analytics-3.png}
\caption{Code: Test.ManAhl.Core.Analytics - page 3}
\label{fig:test.a3}
\end{figure}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{img/code-test-cli.png}
\caption{Code: Test.ManAhl.CLI}
\label{fig:test.cli}
\end{figure}
\end{document}
